---
title: "Forecasting food"
author: "Nickolas Tran and Rim Wolde"
date: "2023-02-22"
output: html_document
runtime: shiny
---
<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #d0e3f1;
}
body {
    background-color:#d0e3f1;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load data, include=FALSE}
library(readxl)
CPIForecastData <- read_excel("CPIForecast.xlsx",
                              skip = 1)
CPIHistoricalData <- read_excel("historicalcpi.xlsx", 
                                skip = 1)
```

## Bar graph

This graph is showing the forecast data for the year 2022 and the graph is showing the overall percentage changes for each food item.

``` {r bar graph, echo=FALSE}

  library(ggplot2)

CPIForecastData_cropped <- head(CPIForecastData, 27)

# Replace non-numeric values with missing values in the "Annual 2022" and Consumer Price Index item" column

CPIForecastData_cropped$`Annual 2022` <- ifelse(grepl("[^0-9.]+", CPIForecastData_cropped$`Annual 2022`),
                                        NA,
                                        as.numeric(gsub("[^0-9.]+", "", CPIForecastData$`Annual 2022`)))

### Subset the data to remove rows with missing values in the "Annual 2022" and "Consumer Price Index item" column

CPIForecastData_complete <- CPIForecastData_cropped[complete.cases(CPIForecastData_cropped$`Annual 2022`), ]
CPIForecastData_complete <- CPIForecastData_cropped[complete.cases(CPIForecastData_cropped$`Consumer Price Index item`), ]

### Bar graph for 2022 Forecast Data

ggplot(data = CPIForecastData_complete, aes(x=`Consumer Price Index item`, y=`Annual 2022`)) +
  geom_bar(stat="identity", fill="steelblue") +
  labs(title="2022 Forecasted percent change of food prices", x="Food Item", y="Percent change") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## Interactive graph

This interactive graph is showing the percentage changes of the price index for beef and veal, eggs and pork throughout decades. 


``` {r interactive line graph with dropbox, message=FALSE}
library(shiny)
library(plotly)
library(readxl)
library(tidyr)

# Read in CPI data
CPIHistoricalData <- read_excel("historicalcpi.xlsx", skip = 1)

# Subset the data to only include meat, eggs, and chicken
meat_eggs_pork <- CPIHistoricalData[CPIHistoricalData$`Consumer Price Index item` %in% c("Beef and veal", "Eggs", "Pork"),]

# Transform the data into long format
meat_eggs_pork <- gather(meat_eggs_pork, key = "Year", value = "Price_Index", -c(`Consumer Price Index item`))

meat_eggs_pork$Price_Index <- as.numeric(meat_eggs_pork$Price_Index)

# Define color palette for each category in the Consumer Price Index item column
colors <- c("Beef and veal" = "firebrick3", "Eggs" = "darkcyan", "Pork" = "darkseagreen3")

# Define UI
ui <- fluidPage(
  
  # Add dropdown for selecting the meat type
  selectInput("meat_type", label = "Select meat type:", 
              choices = c("Beef and veal", "Eggs", "Pork"),
              selected = "Beef and veal"),
  
    # Add slider for selecting the year range
  sliderInput("year_range", label = "Select year range:",
              min = min(as.numeric(meat_eggs_pork$Year)), 
              max = max(as.numeric(meat_eggs_pork$Year)), 
              value = c(min(as.numeric(meat_eggs_pork$Year)), max(as.numeric(meat_eggs_pork$Year))),
              step = 1),
  
  
  # Add plotly output
  plotlyOutput("plot")
)

# Define server
server <- function(input, output) {
  
  # Filter data based on selected meat type
  filtered_data <- reactive({
    meat_eggs_pork[meat_eggs_pork$`Consumer Price Index item` == input$meat_type,]
  })
  

  # Create plotly object
  output$plot <- renderPlotly({
    p <- plot_ly(data = filtered_data(), x = ~Year, y = ~Price_Index, color = ~`Consumer Price Index item`, colors = colors, type = "scatter", mode = "lines+markers") 
    
    # Customize layout
    p <- layout(p, title = "CPI for Meat, Eggs, and Pork", 
                xaxis = list(title = "Year"), 
                yaxis = list(title = "Price Index"))
    
    # Return plot
    p
  })
}

# Run app
shinyApp(ui, server)
```

